Nail: A practical interface generator for data formats

### 2014-04-03 paper/design.tex

A central design decision of Nail is that there is a \emph{semantic bijection} between
the model exposed to the programmer and the byte-level input and output, up to
syntactic equivalence for unambiguous grammars. More precisely, the parser is
the generator's inverse, so parsing the generator's output will yield the
generator's input, but generating the parsers output does not necessarily yield
the parsers input. To understand why this makes sense, consider a grammar for a
text language that tolerates white space, or a binary protocol that tolerates
arbitrarily long padding.\footnote{Say, the physical layer of most communication
  protocols is a possibly infinite sequence of symbols that are syntactically
  nil followed by a pre-determined synchronisation sequence and the actual
  contents of the transmission.} In that case, the program semantics should be
independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded
fields \emph{constants}. Currently, Nail always makes a default choice when
there are multiple options to express a constant, however Nail could be extended
to allow a grammar-specific plug-in to make these choices, say for faster
alignment, streaming applications when data is not ready, or visual appearance in
human-readable protocols. Similarly, Nail does not preserve the layout of
objects referred to by their offsets. If the grammar contains no constants and
offset fields, there is a proper isomorphism between model and protocol.

@@

\paragraph{Fundamental parsers.}
The elementary parsers of Nail are the same as those of Hammer, signed
and unsigned integers with arbitrary lengths up to 64 bits.  Note that is
possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header; in Nail's syntax.  Integer parsers
return their value in the smallest appropriately-sized machine integer
type; e.g., a 24-bit integer is stored in a 32-bit wide variable.

>>>>

A central design decision of Nail is that there is a \emph{semantic bijection} between
the model exposed to the programmer and the byte-level input and output, up to
syntactic equivalence for unambiguous grammars. More precisely, the parser is
the generator's inverse, so parsing the generator's output will yield the
generator's input, but generating the parsers output does not necessarily yield
the parsers input. To understand why this makes sense, consider a grammar for a
text language that tolerates white space, or a binary protocol that tolerates
arbitrarily long padding.\footnote{Say, the physical layer of most communication
  protocols is a possibly infinite sequence of symbols that are syntactically
  nil followed by a pre-determined synchronization sequence and the actual
  contents of the transmission.} In that case, the program semantics should be
independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded
fields \emph{constants}. Currently, Nail always makes a default choice when
there are multiple options to express a constant, however Nail could be extended
to allow a grammar-specific plug-in to make these choices, say for faster
alignment, streaming applications when data is not ready, or visual appearance in
human-readable protocols. Similarly, Nail does not preserve the layout of
objects referred to by their offsets. If the grammar contains no constants and
offset fields, there is a proper isomorphism between model and protocol.

@@

\paragraph{Fundamental parsers.}
The elementary parsers of Nail are the same as those of Hammer, signed
and unsigned integers with arbitrary lengths up to 64 bits.  Note that is
possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header; in Nail's syntax.  Integer parsers
return their value in the smallest appropriately sized machine integer
type; e.g., a 24-bit integer is stored in a 32-bit wide variable.

### 2014-04-03 paper/impl.tex
improve figure placement

@@

Many security vulnerabilities can occur when heap allocations are done
improperly. Therefore, just like Hammer, Nail avoids using the heap as much as
possible, using a custom arena allocator and allocating only fixed-size blocks
from the system allocator (malloc). However, Nail extends upon Hammer's approach and uses two
arenas for each parsed input. One arena is used for intermediate results and is
freed (and zeroed) after parsing completes, whereas the other arena is used only
for allocating the result, and has to be freed by the programmer.

>>>>

\input{fig-utf16}

@@

Many security vulnerabilities can occur when heap allocations are done
improperly. Therefore, just like Hammer, Nail avoids using the heap as much as
possible, using a custom arena allocator and allocating only fixed-size blocks
from the system allocator (malloc). However, Nail extends upon Hammer's approach and uses two
arenas for each parsed input. One arena is used for intermediate results and is
released (and zeroed) after parsing completes, whereas the other arena is used only
for allocating the result, and has to be freed by the programmer.

### 2014-04-03 paper/intro.tex

A promising approach to avoid such vulnerabilities is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesise the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities.  Unfortunately, applying this approach
in practice, using state-of-the-art parser generators, still requires
too much manual programmer effort, making it error-prone, as we describe
next.

>>>>

A promising approach to avoid such vulnerabilities is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities.  Unfortunately, applying this approach
in practice, using state-of-the-art parser generators, still requires
too much manual programmer effort, making it error-prone, as we describe
next.
